---
title: "Automobile Analysis"
Author: "Mahesh"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Info about the data set

Automobile data set consists of three types of entities: (a) the specification of an auto in terms of various characteristics, (b) its assigned insurance risk rating, (c) its normalized losses in use as compared to other cars. The second rating corresponds to the degree to which the auto is more risky than its price indicates. Cars are initially assigned a risk factor symbol associated with its price. Then, if it is more risky (or less), this symbol is adjusted by moving it up (or down) the scale. Actuarians call this process "symboling". A value of +3 indicates that the auto is risky, -3 that it is probably pretty safe. 

The third factor is the relative average loss payment per insured vehicle year. This value is normalized for all autos within a particular size classification (two-door small, station wagons, sports/speciality, etc...), and represents the average loss per car per year. 

Source : https://archive.ics.uci.edu/ml/datasets/automobile


## Reading the data
```{r}
auto_mobiles= read.csv("D://Term I//Automobile_data.csv")
head(auto_mobiles)
library(dplyr)
library(ggplot2)
library(corrplot)
library(reshape)
```


## View the features & thier data types
```{r}
str(auto_mobiles)
```


## Statistical info of the data
```{r}
summary(auto_mobiles)

```


## Inspection for null values
```{r}
c= nrow(auto_mobiles)
nulls=colSums((is.na(auto_mobiles))/c)*100
nulls
```

## But the data set consists of null vlaues with character "?". So apply missing value treatment for those particular columns by replacing "?" with NA

```{r}
#table(auto_mobiles$num.of.doors=='NA')

auto_mobiles[auto_mobiles=='?']= NA



nulls=colSums((is.na(auto_mobiles))/c)*100
nulls


for(i in c(2,19,20,22,23,26))
{
auto_mobiles[,i]=as.character(auto_mobiles[,i])
}


for(i in c(2,19,20,22,23,26))
{
auto_mobiles[,i]=as.numeric(auto_mobiles[,i])
}



for(i in c(2,19,20,22,23,26)){
    auto_mobiles[,i][is.na(auto_mobiles[,i])]<-median(auto_mobiles[,i],na.rm = TRUE)
}



auto_mobiles$make=as.factor(auto_mobiles$make)

nulls=colSums((is.na(auto_mobiles))/c)*100
nulls


## In the above program  '?' is replaced with NA and then NA values are replaced with median of the respective column.
```


### UNIVARIATE ANALYSIS


## Count of vehicles w.r.t  fuel type
```{r}

auto_fuel= auto_mobiles %>% group_by(fuel.type) %>% summarise(No.of.vehicles=n())
ggplot(auto_fuel,aes(fuel.type,No.of.vehicles,fill=fuel.type)) + geom_bar(stat = 'identity',width = 0.6) +theme_bw() +theme_classic() + labs(title="Bar chat @ fuel type")

```

## Histogram of Horse Power
```{r}

hist(auto_mobiles$horsepower)
```

## Count vehicles w.r.t make

```{r}
auto_make= auto_mobiles  %>% group_by(make) %>% summarise(No.of.vehicles= n()) %>% arrange(-No.of.vehicles)
top_10_auto_make= head(auto_make,10)
ggplot(top_10_auto_make,aes(make,No.of.vehicles,fill=make)) + geom_bar(stat = 'identity') +theme_bw() + theme_classic() + labs(title="Bar chat @ make")
```

## Barchart of vehicles based on Aspiration

```{r}
auto_aspiration= auto_mobiles %>%group_by(aspiration) %>% summarise(Count_of_vehicles= n())
ggplot(auto_aspiration,aes(aspiration,Count_of_vehicles,fill=aspiration)) +geom_bar(stat = 'identity',width = 0.5) + theme_bw()+theme_classic()+ labs(title="Bar chat @ Aspiration")
```

## Count of vehicles based on num-of-cylinders

```{r}
auto_cylinders= auto_mobiles %>% group_by(num.of.cylinders) %>% summarise(Count_of_vehicles=n())
ggplot(auto_cylinders,aes(num.of.cylinders,Count_of_vehicles,fill= num.of.cylinders)) + geom_bar(stat = "identity") + theme_bw() + theme_classic() + labs(title="Bar chat @ Num_of_cylinders")
``` 

## Histogram of vehicles with best city-mpg
```{r}
hist(auto_mobiles$city.mpg)
```

## Count of vehicles w.r.t body type

```{r}

auto_door= auto_mobiles %>%  group_by(body.style) %>% summarise(No.of.vehicles=n())
ggplot(auto_door,aes(body.style,No.of.vehicles,fill= body.style)) + geom_bar(stat = "identity") + theme_bw()
```

### BIVARIATE ANALYSIS

## Price vs Make

```{r}

auto_price_make= auto_mobiles %>% group_by(make) %>% summarise(count=n(),Price=max(price)) %>% arrange(-Price)
top_10_manfactures=head(auto_price_make,10)
ggplot(top_10_manfactures,aes(x=reorder(make,-Price),y=Price,fill=make))+ geom_bar(stat = "identity")+ theme_bw()+labs(title="Bar chat @ Top_10_manfactures") 

```

## city mlg vs hwy mlg w.r.t make
```{r}
auto_hvc_make = auto_mobiles[c(3,24,25)]
auto_mlg = melt(auto_hvc_make,id.vars = c("make"))

ggplot(auto_mlg) + geom_bar(aes(x = reorder(make,-value) , fill = variable , y = value), stat = 'identity',position = 'dodge') +theme_bw() + labs(title="BarChart @ City mlg vs Hwy mlg w.r.t make") +theme(axis.text.x = element_text(angle = 90),axis.title.x=element_blank())

```


### Correlation Analysis

```{r}

auto_corr=auto_mobiles[,c(1,2,10,11,12,13,14,17,19,20,21,22,23,24,25)]

auto_correlation=cor(auto_corr)
corrplot(auto_correlation,method = "circle")
```




