---
title: "Time_Series"
author: "TARAH.AI"
date: "24 May 2018"
output: html_document
---

```{r}
library(forecast)
library(dplyr)
```



```{r}
turbo=read.csv("D:/T_Project/Proj_3_Sensor_data/sensor_data.csv",header =T)

colnames(turbo)
turbo_sen = subset(turbo,select= -c(1,4,5,6,28,29))
colnames(turbo_sen)
```



### Plot to determine whether the series is stationary or non-stationary
```{r}
library(ggplot2)

engine1=  turbo_sen %>% filter(engine==1)
dim(engine1)


names=colnames(engine1[,3:23])

for (i in 3:23) {
  
  print(ggplot(engine1, aes(cycles,engine1[i])) + geom_line(color="royalblue") + ylab(colnames(engine1[i]))) 
}

```

## Converting Series to Stationary and then forecasting values for next 5 cycles for first 3 Engines

```{r}
number=1
while(number<=3){
  
  
  turbofilter=turbo_sen%>%filter(engine==number)


for(i in 3:23){
  
  model_forecast=auto.arima(log(turbofilter[,i]))
  forecast_five=forecast(model_forecast,h=5)
  forecast_values=2.718^forecast_five$mean
  print(forecast_values)
  plot(forecast_five,ylab=colnames(turbofilter[i]))
  
}
  number=number+1
}


```


### Transforming Non-Stationary to Stationary series and then forecasting values for Engine_2 
```{r}
turbofilter_1=turbo_sen%>%filter(engine==2)


for(i in 3:17){
  
  model_forecast=auto.arima(log(turbofilter_1[,i]))
  forecast_five=forecast(model_forecast,h=5)
  forecast_value_2= 2.718^forecast_five$mean
  print(forecast_values)
  plot(forecast_five,ylab=colnames(turbofilter_1[i]))
  
}

```




### Transforming Non-Stationary to Stationary series and then forecasting values for Engine_3
```{r}
turbofilter_2=turbo_sen%>%filter(engine==3)


for(i in 3:17){
  
  model_forecast=auto.arima(log(turbofilter_2[,i]))
  forecast_five=forecast(model_forecast,h=5)
  forecast_val= 2.718^forecast_five$mean
  print(forecast_val)
  plot(forecast_five,ylab=colnames(turbofilter_2[i]))
  
}

```


## UDF to forecast the sensor values for desired Engine

```{r}

TS_forecast=function(number){
  
  turbofilter=turbo_sen%>%filter(engine==number)


for(i in 3:23){
  
  model_forecast=auto.arima(log(turbofilter[,i]))
  forecast_five=forecast(model_forecast,h=5)
  forecast_num= 2.718^forecast_five$mean
  plot(forecast_five,ylab=colnames(turbofilter[i]))
  
}
  
}

TS_forecast(1)

```




###################################################################################################################################### FORECASTING USING ARIMA ###########################################
#####################################################################################################


```{r}


filter_sensor=turbo %>% filter(engine==1)

turbo_s2=filter_sensor$sensor_2
turbo_s2=ts(turbo_s2)
plot(turbo_s2)
plot.ts(turbo_s2)

## Determing the values of "p" and "q"

q_value= acf(diff(log(turbo_s2)))

p_value=pacf(diff(log(turbo_s2)))


model1=arima(log(turbo_s2),order = c(1,1,1))
summary(model1) #-2347 AIC
pred1 = forecast(model1,h=5)
pred_1 = 2.718^pred1$mean
pred_1
plot(pred1)

model2=arima(log(turbo_s2),order = c(0,1,1))
summary(model2) # -2348 AIC
pred2 = forecast(model2,h=5)
pred_2 = 2.718^pred2$mean
pred_2
plot(pred2)

model3=arima(log(turbo_s2),order = c(0,1,0))
summary(model3) # -2244 AIC
pred3 = forecast(model3,h=5)
pred_3 = 2.718^pred3$mean
pred_3
plot(pred3)

model4=arima(log(turbo_s2),order=c(1,1,0))
summary(model4) # -2305 AIC
pred4 = forecast(model4,h=5)
pred_4 = 2.718^pred4$mean
pred_4
plot(pred4)
```


## Comparision between ARIMA and Auto ARIMA


```{r}
## Using Auto ARIMA

auto_arima = auto.arima(log(turbo_s2))
summary(auto_arima)

auto_pred = forecast(auto_arima,h=5)
pred_auto = 2.718^auto_pred$mean
pred_auto

plot(auto_pred)


## Using ARIMA
model1=arima(log(turbo_s2),order = c(1,1,1))
summary(model1) #-2347 AIC
pred1 = forecast(model1,h=5)
pred_1 = 2.718^pred1$mean
pred_1
plot(pred1)


## By comparing two models, Auto ARIMA gives the best forecasting values.
```










